---
title: "Movie Recommendation System"
author: "Sam Vuong, Raymond (shanhua) Huang, Carmon Ho, Kyle Murphy"
date: "June 2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

## Overview

In this project, we develop a collaborative filtering recommender (CFR) system for recommending movies. 

If two users share the same interests in the past, e.g. they liked the same movie, they will also have similar tastes in the future.

The collaborative filtering approach considers only user preferences.
It does not take into account the contents of the items (movies) being recommended.

## Libraries used

The following libraries are used:

```{r libs, warning=FALSE, error=FALSE, message=FALSE}
library(arules)
library(caret)
library(dplyr)
library(data.table)
library(ggplot2)
library(ggthemes)
library(lubridate)
library(Matrix)
library(reshape2)
library(recommenderlab)
library(scales)
library(stringr)
library(stringdist)
library(tidyverse)
library(wordcloud)
```

## Dataset

The data has been collected by GroupLens Research and made available rating data sets from the
MovieLens web site (http://movielens.org).

The dataset can be found at: http://files.grouplens.org/datasets/movielens/ml-latest-small.zip

The selected dataset include 4 CSV files:
- movies.csv
- ratings.csv
- tags.csv
- links.csv
Due to time constraints in this project, our recommendation system only uses the rating information
in the ?ratings? data file. It does not use the tag information in the tags data file.

```{r data_load, warning=FALSE, error=FALSE, echo=FALSE}
movies<-read.csv("movies.csv", header=TRUE)
head(movies)

ratings<-read.csv("ratings.csv", header=TRUE)
head(ratings)

tags<-read.csv("tags.csv", header=TRUE)
head(tags)

links<-read.csv("links.csv", header=TRUE)
head(links)
```

## Data Exploration Analysis

```{r data_exploration, warning=FALSE, error=FALSE, echo=FALSE}
# Checking data types of data
str(movies)
str(ratings)
str(tags)
str(links)

# Checking number of movies
length(unique(movies$movieId))
length(unique(ratings$movieId))

# Checking number of users
length(unique(ratings$userId))

# Checking Summary of data
summary(movies)
summary(ratings)

# Convert timestamp to date
ratings <- mutate(ratings, year = year(as_datetime(timestamp, origin="1970-01-01")))
head(ratings)

# Joining Ratings and Movies dataframes
moviesratings <- left_join(ratings, movies, by="movieId")
head(moviesratings)
str(moviesratings)

# Double checking number of movies
length(unique(movies$movieId))
length(unique(ratings$movieId))
length(unique(moviesratings$movieId))

# Checking Summary of data
summary(moviesratings)
```

## Data Visualization

```{r data_visualization, warning=FALSE, error=FALSE, echo=FALSE}
#1. Users vs. Ratings Distribution
moviesratings %>% group_by(userId) %>% summarise(n=n()) %>%
  ggplot(aes(n)) +
  geom_histogram(color = "white") +
  scale_x_log10() + 
  ggtitle("Distribution of Users") +
  xlab("Number of Ratings") +
  ylab("Number of Users") + 
  scale_y_continuous(labels = comma) + 
  theme_economist()

#2. Rating Distribution Histogram
moviercount<-moviesratings %>% group_by(rating) %>% summarise(count=n())
#view(moviercount)
class(moviercount)
moviercount<-as.data.frame(moviercount)

ggplot(moviercount,aes(x=rating,y=count/1000)) +geom_bar(stat="identity") + 
  ggtitle("Rating Distribution") + 
  xlab("Rating") +
  ylab("# Ratings 1*10^3") +
  theme_economist()

#3. Rating vs. Year Distribution
moviesratings %>%
  ggplot(aes(x=year)) +
  geom_bar() +
  xlab("Year") +
  ylab("Number of Ratings") +
  scale_y_continuous(labels = comma) + 
  ggtitle("Rating Distribution Per Year") +
  theme_economist()

#4. Movie Genres Tag Cloud
tag_data <- moviesratings %>% select(genres) %>% 
  group_by(genres) %>% summarise(count = n()) %>% 
  arrange(desc(count)) %>% as.data.frame()

set.seed(1234)
wordcloud(words = moviesratings$genres,
          freq = tag_data$count, min.freq = 1,
          max.words=30, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

#5. vertical bar chart for tags
top_tags <- tags %>%
  group_by(tag) %>%
  summarise(tag_count = n()) %>%
  arrange(desc(tag_count)) %>%
  slice(1:10)

ggplot(top_tags,aes(x=tag,y=tag_count,fill=tag)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  ggtitle("Genre Distribution") + 
  theme(legend.position = "none")
```

## Data Preparation for Modeling

Create a dataframe with only 3 columns required for the Recommender model

Convert it as a matrix, then to realRatingMatrix
Using acast to convert above data as follows:
?????? m1? m2?? m3?? m4
 u1??? 3?? 4??? 2??? 5
 u2??? 1?? 6??? 5
 u3??? 4?? 4??? 2??? 5


```{r model_data_prep, warning=FALSE, error=FALSE, echo=FALSE}
rating_df <- select(moviesratings, userId, movieId, rating)
str(rating_df)
head(rating_df)

rating_df_matrix<-acast(rating_df, userId ~ movieId, value.var = "rating")
class(rating_df_matrix)

# Convert it into matrix
rating_matrix_raw<-as.matrix(rating_df_matrix)
class(rating_matrix_raw)

# Convert it into realRatingMatrix data structure
# realRatingMatrix is a recommenderlab sparse-matrix like data-structure
rating_matrix <- as(rating_matrix_raw, "realRatingMatrix")
rating_matrix

# view rating_matrix in other possible ways
#as(rating_matrix, "list")???? # A list
#as(rating_matrix, "matrix")?? # A sparse matrix
head(as(rating_matrix, "data.frame")) # View as a data-frame

# normalize the rating matrix
rating_matrix_norm <- normalize(rating_matrix)
rating_matrix_norm
head(as(rating_matrix_norm, "data.frame")) # View as a data-frame

# Can also turn the matrix into a 0-1 binary matrix
rating_matrix_bin <- binarize(rating_matrix, minRating=2)
rating_matrix_bin
head(as(rating_matrix_bin, "data.frame")) # View as a data-frame
```

## Visualize the ratings matrix

Draw an image plot of raw-ratings & normalized ratings
? A column represents one specific movie and ratings by users
?? are shaded.
?? Note that some items are always rated 'black' by most users
??? while some items are not rated by many users
???? On the other hand a few users always give high ratings
????? as in some cases a series of black dots cut across items

```{r visualize_matrix, warning=FALSE, error=FALSE, echo=FALSE}
#image(rating_matrix,      main = "Raw Ratings")?????? 
#image(rating_matrix_norm, main = "Normalized Ratings")
```

## Create Recommender model

```{r rec_model, warning=FALSE, error=FALSE, echo=FALSE}
# Create a Recommender model using UBCF (user-based collaborative filtering)
rec=Recommender(rating_matrix,method="UBCF",param=list(normalize="Z-score",method="Cosine", nn=5))

# Examine the model we got
print(rec)
getModel(rec)$data
getModel(rec)$nn
names(getModel(rec))
```

## Create predictions

```{r rec_predict, warning=FALSE, error=FALSE, echo=FALSE}
recom <- predict(rec, rating_matrix, type="ratings")
recom

# Convert all recommendations to list structure. 
rec_list<-as(recom,"list")
head(summary(rec_list))
```

## Check movie recommendations with n users in test data

```{r rec_check, warning=FALSE, error=FALSE, echo=FALSE}
n_users  <- 5
n_movies <- 10
# For the users in test file, one by one
for ( u in 1:n_users)
{
  # Get userid
  uid  <- u

  # Obtain top n recommendations for the user
  recom <- predict(rec, rating_matrix[uid], n=n_movies)
  # Convert it to readable list
  recom_list <- as(recom, "list")  

  # Obtain movie titles of recommendations from movies dataset
  rec_movies <- matrix(0,n_movies)
  for (i in 1:n_movies){
    rec_movies[i] <- as.character(subset(movies,
                                  movies$movieId == as.integer(recom_list[[1]][i]))$title)
  }
  print(paste("==== Movie Recommendations For User: ", uid, " ===="))
  print(rec_movies)
  print("")
}
```

## Evaluation

```{r rec_evaluation, warning=FALSE, error=FALSE, echo=FALSE}
# Take subset of data users with more than 50 ratings
rating_matrix <- rating_matrix[rowCounts(rating_matrix) >50,]
rating_matrix
```

##  Create an Evaluation Scheme

using train/test 80/20 split validation
with given=5 means with 5 items given scheme
with goodRating=3 means items with user rating >= 3 are considered positives

```{r eval_scheme, warning=FALSE, error=FALSE, echo=FALSE}
eval_scheme <- evaluationScheme(rating_matrix,
                                method="split",train=0.80,k=1,given=5,goodRating=3) 
eval_scheme
```

## Perform Evaluation

Compare UBCF algorithm other basic algorithms
Here we compare the UBCF model (user-based collaborative filtering)
to the POPULAR model (based on item popularity)
and the RANDOM model (random recommendations)

```{r eval_topNList, warning=FALSE, error=FALSE, echo=FALSE}
algorithms <- list(
  UBCF = list(name = "UBCF", param = NULL),
  POPULAR = list(name = "POPULAR", param = NULL),
  RANDOM  = list(name = "RANDOM",  param = NULL)
)

# Evaluate using top-N recommendation lists
eval_topNList <- evaluate(eval_scheme,
                         algorithms,type="topNList",n=c(1,3,5,10,15,20))
# Print results
avg(eval_topNList)

# Plot ROC Curse
plot(eval_topNList, annotate=TRUE, main="ROC Curve")

plot(eval_topNList, "prec/rec", annotate=TRUE, main="Precision-Recall")
```

```{r eval_ratings, warning=FALSE, error=FALSE, echo=FALSE}
# Evaluate prediction of missing ratings
eval_ratings <- evaluate(eval_scheme, algorithms, type="ratings")

# Print results
avg(eval_ratings)

# Print results
plot(eval_ratings)
```